---
title: "Toyota-Stanford Replication"
output:
  pdf_document:
    toc: true
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

Author: Brenden Eum (2025)

# Preamble

Things you can adjust.
```{r}
#rm(list=ls())
seed = 4
```


Start fresh.
```{r}
library(tidyverse)
library(here) # Looks for ".here" file and treats that as the root folder! Great for project management!
library(lme4)
library(sandwich)
library(plotrix)
library(patchwork)
library(car)
library(corrplot)

.root <- here()
.outdir = file.path(.root, "outputs")
.datadir = file.path(.root, "data")
.tempdir = file.path(.root, "analysis/temp")
.figdir = file.path(.root, "analysis/figures")
.tabdir = file.path(.root, "analysis/tables")
.txtdir = file.path(.root, "analysis/text")
```

Quarterly sales data from autodealerdata
```{r}
qs = read.csv(file.path(.datadir, "quarterly_sales_2005_2025.csv"))
```

NAcc data from Stanford
- subject: subject id
- Make: brand of car
- Model: car model
- nacc: % signal change at nacc8mm_tr4
- mpfc: % signal change at mpfc8mm_tr5, tr6, tr7
- insula_hypergranular_ana_tr4: % signal change at posterior insula, which we may use as a control for noise
- knowMore_clean: ratings (1-4) how much the subjects wanted to know more
- purchase_clean: ratings (1-4) subject's purchase likelihood

```{r}
nacc = read.csv(file.path(.datadir, "stanford-nacc/Car2_agg_df_by_subject_run2 (1).csv"))
nacc = nacc %>% 
  mutate(
    make = Make %>% tolower(),
    make = ifelse(make=="mercedes", "mercedes-benz", make),
    model = Model %>% tolower() %>% gsub(" ", "", .),
    year = 2023,
    pins = insula_hypergranular_ana_tr4,
    electric = ifelse(Fuel_Type.x=="Electric", 1, 0)
  ) %>%
  select(
    subject, year, make, model, electric, q_slope_2023,
    nacc, mpfc, ains, pins, 
    knowMore_clean, purchase_clean
  )
```


Get all the stimuli used in the fMRI study.
```{r}
nacc_stim = nacc %>%
  distinct(year, make, model) %>%
  print()
nStim = nrow(nacc_stim)
```

# Analysis Explained

## Problem

Goodcarbadcar dataset does not distinguish different manufacturing dates for each make-model in their sales counts. Therefore, sales in 2023-Q3 could reflect both 2023 and 2024 models. 

Autodealerdata dataset DOES distinguish different manufacturing dates for each make-model in their sales counts.

## Solutions

We will try two approaches. Either goodcarbadcar sales counts reflects (1) sales of the newest make-model, or (2) cumulative sales of that make-model. 

Typically, dealerships do not keep cars on a lot for more than 18 months. This means sales counts in a specific quarter will, at most, reflect the newest model and the previous year's model. 

Solution 1: Newest Model

For each stimulus, set quarterly sales equal to:
- Q1: 2023-Q1 sales for 2023 model
- Q2: 2023-Q2 sales for 2023 model
- Q3.1: 2023-Q3 sales for 2024 model
- Q4.1: 2023-Q4 sales for 2024 model

Solution 2: Oldest Model

For each stimulus, set quarterly sales equal to:
- Q1: 2023-Q1 sales for 2023 model
- Q2: 2023-Q2 sales for 2023 model
- Q3: 2023-Q3 sales for 2023 model
- Q4: 2023-Q4 sales for 2023 model

Solution 3: Both Models

For each stimulus, set quarterly sales equal to:
- Q1: 2023-Q1 sales for 2023 model
- Q2: 2023-Q2 sales for 2023 model
- Q3: 2023-Q3 sales for 2023 model + 2023-Q3 sales for 2024 model
- Q4: 2023-Q4 sales for 2023 model + 2023-Q4 sales for 2024 model

Then, use these metrics to calculate the average change in quarterly sales over the year.
dQ = (1/3)[(Q2-Q1) + (Q3-Q2) + (Q4-Q3)] 


### Solution 1: Newest Model

```{r}
df_dQ_New = data.frame()

for (s in 1:nStim) {
  
  make = nacc_stim$make[s]
  model = nacc_stim$model[s]
  
  ind1 = (qs$year==2023 & qs$make==make & qs$model==model) & (qs$year_sold==2023 & qs$quarter_sold==1)
  ind2 = (qs$year==2023 & qs$make==make & qs$model==model) & (qs$year_sold==2023 & qs$quarter_sold==2)
  ind3 = (qs$year==2024 & qs$make==make & qs$model==model) & (qs$year_sold==2023 & qs$quarter_sold==3)
  ind4 = (qs$year==2024 & qs$make==make & qs$model==model) & (qs$year_sold==2023 & qs$quarter_sold==4)
  
  if (sum(ind1)!=0 && sum(ind2)!=0 && sum(ind3)!=0 && sum(ind4)!=0) {
    
    Q1 = qs[ind1, "count"]
    Q2 = qs[ind2, "count"]
    Q3 = qs[ind3, "count"]
    Q4 = qs[ind4, "count"]
    
    dQ_New = (1/3)*((Q2-Q1)+(Q3-Q2)+(Q4-Q3))
  
    add_row = data.frame(make, model, dQ_New)
    df_dQ_New = bind_rows(df_dQ_New, add_row)
    
  }
}

epsilon = 0.0001
df_dQ_New$ndQ_New = (df_dQ_New$dQ_New + abs(min(df_dQ_New$dQ_New)) + epsilon) %>% log()
```


### Solution 2: Oldest Model

```{r}
df_dQ_Old = data.frame()

for (s in 1:nStim) {
  
  make = nacc_stim$make[s]
  model = nacc_stim$model[s]
  
  ind1 = (qs$year==2023 & qs$make==make & qs$model==model) & (qs$year_sold==2023 & qs$quarter_sold==1)
  ind2 = (qs$year==2023 & qs$make==make & qs$model==model) & (qs$year_sold==2023 & qs$quarter_sold==2)
  ind3 = (qs$year==2023 & qs$make==make & qs$model==model) & (qs$year_sold==2023 & qs$quarter_sold==3)
  ind4 = (qs$year==2023 & qs$make==make & qs$model==model) & (qs$year_sold==2023 & qs$quarter_sold==4)
  
  if (sum(ind1)!=0 && sum(ind2)!=0 && sum(ind3)!=0 && sum(ind4)!=0) {
    
    Q1 = qs[ind1, "count"]
    Q2 = qs[ind2, "count"]
    Q3 = qs[ind3, "count"]
    Q4 = qs[ind4, "count"]
    
    dQ_Old = (1/3)*((Q2-Q1)+(Q3-Q2)+(Q4-Q3))
    Q_Old = Q1 + Q2 + Q3 + Q4
  
    add_row = data.frame(make, model, dQ_Old, Q_Old)
    df_dQ_Old = bind_rows(df_dQ_Old, add_row)
    
  }
}

epsilon = 0.0001
df_dQ_Old$ndQ_Old = (df_dQ_Old$dQ_Old + abs(min(df_dQ_Old$dQ_Old)) + epsilon) %>% log()
```


### Solution 3: Both Models

```{r}
df_dQ_Both = data.frame()

for (s in 1:nStim) {
  
  make = nacc_stim$make[s]
  model = nacc_stim$model[s]
  
  ind1 = (qs$year==2023 & qs$make==make & qs$model==model) & (qs$year_sold==2023 & qs$quarter_sold==1)
  ind2 = (qs$year==2023 & qs$make==make & qs$model==model) & (qs$year_sold==2023 & qs$quarter_sold==2)
  ind3.1 = (qs$year==2023 & qs$make==make & qs$model==model) & (qs$year_sold==2023 & qs$quarter_sold==3)
  ind3.2 = (qs$year==2024 & qs$make==make & qs$model==model) & (qs$year_sold==2023 & qs$quarter_sold==3)
  ind4.1 = (qs$year==2023 & qs$make==make & qs$model==model) & (qs$year_sold==2023 & qs$quarter_sold==4)
  ind4.2 = (qs$year==2024 & qs$make==make & qs$model==model) & (qs$year_sold==2023 & qs$quarter_sold==4)
  
  if (sum(ind1)!=0 && sum(ind2)!=0 && sum(c(ind3.1,ind3.2))!=0 && sum(c(ind4.1,ind4.2))!=0) {
    
    Q1 = qs[ind1, "count"]
    Q2 = qs[ind2, "count"]
    Q3.1 =  if (sum(ind3.1)!=0) qs[ind3.1, "count"] else 0
    Q3.2 =  if (sum(ind3.2)!=0) qs[ind3.2, "count"] else 0
    Q3 = Q3.1 + Q3.2
    Q4.1 =  if (sum(ind4.1)!=0) qs[ind4.1, "count"] else 0
    Q4.2 =  if (sum(ind4.2)!=0) qs[ind4.2, "count"] else 0
    Q4 = Q4.1 + Q4.2
    
    dQ_Both = (1/3)*((Q2-Q1)+(Q3-Q2)+(Q4-Q3))
  
    add_row = data.frame(make, model, dQ_Both)
    df_dQ_Both = bind_rows(df_dQ_Both, add_row)
    
  }
}

epsilon = 0.0001
df_dQ_Both$ndQ_Both = (df_dQ_Both$dQ_Both + abs(min(df_dQ_Both$dQ_Both)) + epsilon) %>% log()
```


# Individual-Level Replications

## Make Regression Data

```{r}
# Merge in autodealerdata average quarterly slopes
rdata = nacc %>%
  merge(., df_dQ_New, by=c("make", "model"), all.x=T) %>%
  merge(., df_dQ_Old, by=c("make", "model"), all.x=T) %>%
  merge(., df_dQ_Both, by=c("make", "model"), all.x=T)

hist_dQ_New = ggplot(data=rdata) +
  geom_histogram(aes(x=(ndQ_New))) +
  theme_classic()
hist_dQ_Old = ggplot(data=rdata) +
  geom_histogram(aes(x=(ndQ_Old))) + 
  theme_classic()
hist_dQ_Both = ggplot(data=rdata) +
  geom_histogram(aes(x=(ndQ_Both))) + 
  theme_classic()
hist_dQ_New / hist_dQ_Old / hist_dQ_Both
```

## Check corr between datasets

goodcarbadcar data vs autodealer data.
```{r}
corrdata = rdata %>% distinct(make, model, .keep_all=T)
ggplot(data=corrdata, aes(y=q_slope_2023)) +
  geom_abline(intercept=0, slope=1, color="grey", linewidth=2) +
  
  geom_line(aes(x=dQ_Old, color="Old"), linewidth=1, alpha=.8) +
  geom_line(aes(x=dQ_New, color="New"), linewidth=1, alpha=.8) +
  geom_line(aes(x=dQ_Both, color="Both"), linewidth=1, alpha=.8) +
  
  labs(
    y = "goodcarbadcar dQ",
    x = "autodealerdata dQ"
  ) +
  theme_bw() +
  scale_color_manual(
    name="Manufacture Year", 
    values=c("Old"="dodgerblue", "New"="gold2", "Both"="maroon")
  )
```
```{r}
lm(q_slope_2023 ~ dQ_Old, data = corrdata) %>% summary()
```

```{r}
cor.test(corrdata$q_slope_2023, corrdata$dQ_Old)
```



## Replication 1: Newest

```{r}
# replication_1 <- lmer(
#   dQ_New ~ 1 + nacc + (1 + nacc | subject), 
#   data = rdata
# )
# 
# summary(replication_1)
```


## Replication 2: Oldest

Anova analysis suggests that linear model (replication_2.2) is performing better than random effects model (replication_2.1).
```{r}
replication_2.1 <- lmer(
  dQ_Old ~ 1 + electric + knowMore_clean + nacc + mpfc + pins + (1 + nacc | subject), 
  data = rdata,
  control = lmerControl(optimizer = "bobyqa")
)

replication_2.2 <- lm(
  dQ_Old ~ 1 + electric + knowMore_clean  + nacc + mpfc + pins, 
  data = rdata
)

anova(replication_2.1, replication_2.2)
```

```{r}
replication_2.2_noBrain <- lm(
  dQ_Old ~ 1 + electric + knowMore_clean, 
  data = rdata
)
```

```{r}
replication_2.2_noBrain %>% summary()
```

```{r}
summary(replication_2.2)
```

What about with goodcarbadcar data?
```{r}
replication_2.3 <- lm(
  q_slope_2023 ~ 1 + electric + knowMore_clean  + nacc + mpfc + ains, 
  data = rdata
)
summary(replication_2.3)
```


Average change in quarterly sales, by average NAcc signal.
```{r}
pdata = rdata %>%
  group_by(make, model) %>%
  summarize(
    nacc_se = std.error(nacc),
    nacc_y = mean(nacc),
    mpfc_y = mean(mpfc),
    dQ_Old = first(dQ_Old),
    dQ_New = first(dQ_New),
    dQ_Both = first(dQ_Both),
    electric = first(electric)
  )
pdata$electric = factor(pdata$electric, levels=c(0,1), labels=c("Gas", "Electric"))

corr_res = cor.test(pdata$dQ_Old, pdata$nacc_y)
corr_est = corr_res$estimate %>% round(2) %>% as.character()
corr_pval = corr_res$p.value %>% round(2) %>% as.character()

ggplot(data=pdata) +
  
  geom_smooth(
    aes(y=dQ_Old, x=nacc_y), method="lm", se=TRUE, 
    color="grey70", fill="grey80", linetype="dashed"
  ) +
  geom_point(aes(y=dQ_Old, x=nacc_y, color=electric)) +
  geom_text(
    aes(y=dQ_Old, x=nacc_y, label=model), 
    hjust=0, vjust=0, nudge_x=-.002, nudge_y=220, size=3
  ) +
  annotate(
    "text", x=Inf, y=-Inf, hjust=1.1, vjust=-1.1, 
    label=paste0("R = ", corr_est, ", p = ", corr_pval), 
    size=4
  ) +
  
  labs(
    y = "Avg. Quarterly Slope in 2023",
    x = "NAcc (% Signal Change, tr4)",
    color = "Fuel Type"
  ) +
  theme_bw()
```

```{r}
cor.test(rdata$knowMore_clean, rdata$nacc)
```

```{r}
cor.test(rdata$purchase_clean, rdata$nacc)
```

VIF Analysis
```{r}
# Calculate VIF for replication_2.2 model
vif_results <- vif(replication_2.2)

# Create a data frame for better visualization
vif_df <- data.frame(
  Variable = names(vif_results),
  VIF = as.numeric(vif_results)
)

# Add interpretation column
vif_df$Interpretation <- ifelse(vif_df$VIF > 10, "High multicollinearity (VIF > 10)",
                       ifelse(vif_df$VIF > 5, "Moderate multicollinearity (5 < VIF < 10)",
                              "Low multicollinearity (VIF < 5)"))
print(vif_df)

# Check correlation matrix of predictors for additional insight
predictor_vars <- c("electric", "knowMore_clean", "nacc", "mpfc", "ains", "pins")
correlation_matrix <- cor(rdata[, predictor_vars], use = "complete.obs")

# Visualize correlation matrix
corrplot(correlation_matrix, 
         method = "color", 
         type = "upper", 
         order = "hclust",
         tl.cex = 0.8,
         tl.col = "black",
         title = "Correlation Matrix of Predictors",
         mar = c(0,0,1,0))
```




## Replication 3: Both

Anova analysis suggests that linear model (replication_3.2) is performing better than random effects model (replication_3.1).
```{r}
replication_3.1 <- lmer(
  dQ_Both ~ 1 + electric + knowMore_clean + nacc + mpfc + ains + (1 + nacc | subject), 
  data = rdata,
  control = lmerControl(optimizer = "bobyqa")
)

replication_3.2 <- lm(
  dQ_Both ~ 1 + electric + knowMore_clean  + nacc + mpfc + ains, 
  data = rdata
)

anova(replication_3.1, replication_3.2)
```

```{r}
summary(replication_3.2)
```


# Extra

## Raw Yearly Sales

```{r}
lm(log(Q_Old) ~ 1 + scale(electric) + scale(knowMore_clean) + scale(nacc) + scale(mpfc) + scale(ains), data=rdata) %>% summary()
```

```{r}
lm(log(Q_Old) ~ 1 + scale(electric) + scale(knowMore_clean), data=rdata) %>% summary()
```

```{r}
lm(scale(q_slope_2023) ~ 1 + scale(electric) + scale(knowMore_clean) + scale(nacc) + scale(mpfc) + scale(ains), data=rdata) %>% summary()
```

```{r}
lm(scale(q_slope_2023) ~ 1 + scale(electric) + scale(knowMore_clean), data=rdata) %>% summary()
```
