---
title: "07_check_autodealerdata_consistentDealers"
output:
  pdf_document:
    toc: true
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

Checking validity of autodealerdata, when subsetted to consistent dealers between 2017 to 2024.

## Start Fresh
```{r}
rm(list=ls())
gc()
library(here)
library(tidyverse)
library(arrow)
library(plotrix)
library(patchwork)
library(lubridate)
.root = here() 
.outdir = file.path(.root, "outputs")
.datadir = file.path(.root, "data")
.tmpdir = file.path(.root, "analysis/temp")
.figdir = file.path(.root, "analysis/figures")
.tabdir = file.path(.root, "analysis/tables")
.txtdir = file.path(.root, "analysis/text")
```

## Subset by Consistent Dealers

Only keep dealers that are consistently in the dataset between 2017 and 2024

```{r}
cache_file = file.path(.tmpdir, "autodealerdata2_individual_parquets")
individual_sales = open_dataset(cache_file)

# Keep only observations from consistent dealers in the 2017-2024 window
individual_new_sales = individual_sales %>%
  filter(is_new == T & year > 2005 & year < 2025) %>%
  filter(year_sold >= 2017 & year_sold <= 2024) %>% 
  collect()

# Check how many dealers were consistent and how many observations remain
#cat("Number of consistent dealers:", length(consistent_dealers), "\n")
cat("Observations in filtered dataset:", nrow(individual_new_sales), "\n")
```


# Where are these dealers?

Note, we're only considering zip codes that have 5 digits.

```{r}
library(zipcodeR)
library(maps)
library(viridis)

# Get counts per zipcode
zips = individual_new_sales$zip_code[individual_new_sales$zip_code>=10000 & individual_new_sales$zip_code<=99999]
zips = data.frame(zipcode=as.character(zips))
zips = zips %>% group_by(zipcode) %>% summarize(count=n(), .groups="drop")

# Get zip code data with coordinates
zip_data <- zip_code_db %>%
  inner_join(zips, by = "zipcode") %>%
  select(zipcode, lat, lng, count)

# Get US map data
us_map <- map_data("state")

# Create the heatmap with log scaling
ggplot() +
  geom_polygon(data = us_map, 
               aes(x = long, y = lat, group = group),
               fill = "white", color = "gray50", size = 0.2) +
  geom_point(data = zip_data,
             aes(x = lng, y = lat, color = count),
             alpha = 0.8, size=0.2) +
  scale_color_gradient(name = "Count", 
                      low = "mistyrose", 
                      high = "dodgerblue3",
                      trans = "log10",
                      labels = scales::comma_format(),
                      limits = c(1, 1000000)) +
  coord_fixed(1.3) +
  theme_void() +
  theme(legend.position = "right",
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

# Unique dealers by year

```{r}
unique_dealers_by_year = individual_new_sales %>%
  group_by(year_sold, dealer_id) %>%
  summarize(make = first(model), .groups="drop") %>%
  group_by(year_sold) %>%
  summarize(count = n(), .groups="drop")

ggplot(data=unique_dealers_by_year, aes(x=year_sold, y=count)) +
  geom_line(linewidth=1) +
  labs(y="# of Unique Dealers", x="Year Sold") +
  coord_cartesian(ylim=c(0, NA)) +
  theme_bw()
```




# Transactions Per Year

```{r}
#individual_new_sales = read_parquet(file.path(.tmpdir, "individual_new_sales_2005_2025.parquet"))

FRED = read.csv("/Users/brenden/Desktop/motorVAE/data/FRED_lightweight_vehicle_monthly_sales.csv") %>%
  mutate(
    year_sold = year(as.Date(observation_date))
  ) %>%
  group_by(year_sold) %>%
  summarize(
    total_sales = sum(ALTSALES, na.rm = TRUE),
    n_months = n()  # Count months to verify complete years
  ) 

df_trans_year = individual_new_sales %>%
  group_by(year_sold) %>%
  summarize(count = n(), .groups="drop")

# Calculate scaling factor for the secondary axis
scale_factor <- max(df_trans_year$count) / max(FRED$total_sales)

# Plot
ggplot() +
  
  geom_line(
    data=df_trans_year, 
    aes(x=year_sold, y=count, color="Autodealerdata, Consistent Dealers"),
    linewidth=1
  ) +
  
  geom_line(
    data=FRED, 
    aes(x=year_sold, y=total_sales*scale_factor, color="FRED, Lightweight Vehicle Sales"),
    linewidth=1
  ) +
  
  scale_y_continuous(
    name = "Autodealerdata New Car Sales",
    sec.axis = sec_axis(~ . / scale_factor, name = "FRED Lightweight Vehicle Sales (millions)")
  ) +
  # Add legend
  scale_color_manual(values = c(
    "Autodealerdata, Consistent Dealers" = "dodgerblue",
    "FRED, Lightweight Vehicle Sales" = "maroon")
  ) +
  labs(x="Year Sold", color="Data Source") +
  theme_bw() +
  theme(legend.position="bottom") 
```

```{r}
cor(df_trans_year$count, FRED$total_sales[FRED$year_sold<2025])
```



# Unique Makes & Models Per Year

```{r}
#individual_new_sales = read_parquet(file.path(.tmpdir, "individual_new_sales_2005_2025.parquet"))

df_models_year = individual_new_sales %>%
  distinct(year_sold, make, model) %>%
  group_by(year_sold) %>%
  summarize(count=n(), .groups="drop")

ggplot(data=df_models_year, aes(x=year_sold, y=count)) +
  geom_line(linewidth=1) +
  labs(y="# of Unique Make-Models of New Cars", x="Year Sold") +
  coord_cartesian(ylim=c(0, NA)) +
  theme_bw()
```


# Mean Asking Price by Year

```{r}
#individual_new_sales = read_parquet(file.path(.tmpdir, "individual_new_sales_2005_2025.parquet"))

df_ask_year = individual_new_sales %>%
  group_by(year_sold) %>%
  summarize(ask_mean = mean(ask_price), ask_se = std.error(ask_price), ask_median = median(ask_price), .groups="drop")

ggplot(data=df_ask_year, aes(x=year_sold)) +
  geom_line(aes(y=ask_median, color="Median"), linewidth=1) +
  geom_line(aes(y=ask_mean, color="Mean"), linewidth=1) +
  geom_linerange(aes(ymin=ask_mean-ask_se, ymax=ask_mean+ask_se, color="Mean"), linewidth=1) +
  labs(y="Ask Price", x="Year Sold", color="Statistic") +
  theme_bw()
```


# Sales by Age and Proportion of New by Age

```{r}
rm(individual_new_sales)
gc()

df_new_age = individual_sales %>%
  filter(year > 2005 & year < 2025) %>%
  filter(year_sold >= 2017 & year_sold <= 2024) %>% 
  #filter(dealer_id %in% consistent_dealers) %>%
  mutate(age = year_sold - as.integer(year)) %>%
  group_by(age) %>%
  summarize(new_proportion = mean(is_new), count=n(), .groups="drop") %>%
  collect()

p1 = ggplot(data=df_new_age, aes(x=age, y=count)) +
  geom_linerange(aes(ymin=0, ymax=count)) +
  geom_point() +
  labs(y="Count", x="Age") +
  theme_bw()

p2 = ggplot(data=df_new_age, aes(x=age, y=new_proportion)) +
  geom_hline(yintercept = .5, color="grey") +
  geom_point() +
  geom_linerange(aes(ymin=0, ymax=new_proportion)) +
  labs(y="Pr. Reported as New Vehicle Sold", x="Age of Vehicle") +
  theme_bw()

p1 / p2
```
